---
title: "Test"
author: "Joshua Matthew(SCV2021)"
date: "18/10/2019"
output: html_document
---





Q1; Write R program to create a matrix taking a given vector of numbers as input. Display the matrix.
```{r}
matrix1<-matrix(1:20, nrow = 5, ncol = 4)
matrix1
```

Q2: Write R program to create “two” 2x3 matrix and add, subtract, multiply the matrices. 
```{r}
matrixA<- matrix(5:10, nrow = 2, ncol = 3)
matrixB<- matrix(3:8, nrow =  2, ncol = 3)
matrixAdd<- matrixA + matrixB
matrixSubtract<- matrixA- matrixB
matrixMultiply<- matrixA * matrixB

matrixAdd
matrixSubtract
matrixMultiply
```

Q3: Write R program to access the element at 3rd column and 2nd row of a given matrix.
```{r}
#from matrixA

matrixA[2,3]
```
Q4: Write R program to create vector of numeric, complex, logical and character types of length 6

```{r}
R_numeric<- c(1:6)
R_Logical<- c(FALSE,TRUE,TRUE,TRUE,FALSE,TRUE)
R_Character<- c('a','b','c','d','e','f')
R_Complex<- c('a',1,'e',FALSE,4,TRUE)
R_Complex
R_Character
R_Logical
R_numeric
```


Q5: Write R program to multiply two vectors of integer type of length 3
```{r}
vector1<- c(2,3,5)
vector2<- c(5,2,9)
multiplyVectors<- vector1 * vector2
multiplyVectors
```

Q6:  Load the Tumor data set into R using the read.table() function.

```{r}
tumor<- read.table("data/tumor.txt", header = TRUE, sep = "\t")
view(tumor)
```

Q7: Using the data in Q6, Plot the scatter plot of the variables “time”  and “futime “, with "time" on the y-axis and "futime" on the x-axis.  Format your plot by putting title on the plot.
```{r}
plotTumor<- plot(tumor$futime,tumor$time, main = "plot of futime against time", xlab = "time", ylab = "futime")
plotTumor
```

Q8: Using the data in Q6, Carry out histogram plot of the “time” variable and Title you plot
```{r}

histplotTumor<- hist(tumor$time, main = "Histogram plot of time")
histplotTumor
```
