---
title: "THird session"
author: "Joshua Matthew(SCV2021)"
date: "17/10/2019"
output: pdf_document
---

```{r}
library(tidyverse)
library(readr)
library(readxl)
#library(hflights)

```

Hey!
#Tidyverse functions:
--gather(range of columns to be gathered, key = column where they exist, value = the values they should reference)
--seperate
--spread


we'll be looking into data manipulation and transformation 
import the excel file 'mbta.xlsx' and assign it to a variable 'New_dta'

```{r}
New_dta<-read_excel("data/mbta.xlsx", skip = 1, range =  cell_cols(2:60))

```

```{r}
View(New_dta)
New_dta
```

the gather function transforms/transposes multiple columns into one
% means then. > means next action. 
key is a keyword for renaming the column being transformed and 'value' assigns name to the column holding the 'year's' corresponding values.
```{r}
New_dta_tidy<- New_dta%>%gather( '2007-01' : '2011-10', key = "year", value = "passenger")
```

```{r}
View(New_dta_tidy)
```

```{r}
New_dta_tidy[(1:20),c(1,3)]
```

using the separate function to distinguish year from month in the  gathered table 'New_dta_tidy'
```{r}
New_dta_tidySep<-New_dta_tidy%>%separate(year, into = c("Year","Month"))
```

```{r}
View(New_dta_tidySep)
```

spread is the opposite of gather as it opens a column's rows into columns
the first argument is the column to be spread and the second argument is the column whose value will spread accross the first argument
```{r}
New_dta_tidySpread<- New_dta_tidySep %>%spread(mode, passenger)
```

```{r}
View(New_dta_tidySpread)
```

to now select columns of interest from the spread data
```{r}
Some_cols<-New_dta_tidySpread%>%.[,c(1,2,6,7,8)]
```
cols of interest :year, month, commuter rail, light rail and heavy rail
```{r}
View(Some_cols)
```

gather the last 3 cols into a column named 'rails' against a value 'passenger'
```{r}
Some_cols_gather<-Some_cols%>%gather('Commuter Rail':'Light Rail', key = "all_rails", value = "passenger" )
```

```{r}
view(Some_cols_gather)
#to plot two selected columns
toplot<-Some_cols_gather[1:58,]
plotted<-plot(toplot$Month, toplot$passenger, main = "Months over passengers", xlab = "months", ylab = "passengers")

```

```{r}
library(dplyr)
data("hflights")
view(hflights)

```

DPLYR functions:
--filter
#the dplyr 'filter' function selects the rows that meets the set conditions. eg.  filter(table_name, criteria1 == "criteria value", criteria2 == "criteria2 value")
-- between== 
-- arrange== arrange(table name, column, column) asc by default
-- select==  select(table name, column lits to be view seperated by comma)
            --select has functions which can rename a column but will not pull other unrenamed columns in its result.
            --select can specify 'starts_with("keyword's first letters"))


to use the filter function on hflight data
```{r}
flights<- hflights
filteredFlights<- filter(flights, Year== 2011, Month == 3, Dest == "MIA")
filteredFlights
```

to filter data from the column only when the values in the column meets a criteria.
```{r}
new_filteredDest<- filter(flights, Dest%in%c("FLL", "MIA"))
new_filteredDest
```

to use the 'between' with the filter function function 
filter(tableName, between(ColumnToBeFiltered, left-initial,right-final))[]
```{r}
flightRange<- filter(flights, between(Distance, 224, 944))[2300:2310,11:16]
flightRange
```



exercise: find all flights that 
-a; departed in april 2011
-b; operated by AA and WN
```{r}
View(flights)
```

```{r}
aprilFlight<-filter(flights, Year== 2011, Month==4, UniqueCarrier%in%c("AA","WN"))
view(aprilFlight)
```

to arrange the data set based on columns

```{r}
arrangedData<- arrange(aprilFlight,Year, desc(DepTime))
View(arrangedData)

```

to use the dplyr's select function to select specific columns
```{r}
selected<- dplyr::select(arrangedData, Year,DepTime, Month)[1:10,]
View(selected)
```

select and rename specific columns
```{r}
view(dplyr::select(arrangedData, newYear = Year, newMonth = Month))

```

to rename a set of columns
```{r}
view(rename(arrangedData, newYear = Year, Destination = Dest))
```
to search in the select function
```{r}
view(dplyr::select(arrangedData, starts_with("Dep"), ends_with("arr")))
```
using one_of function with select statement
```{r}
joe<- c("Year","Month","DayOfMonth","ArrTime")
shua<-dplyr:: select(flights, one_of(joe))
head(shua)
```

using the mutate function
```{r}
calc<-flights%>%dplyr::select(ends_with("Delay"),Distance,AirTime)%>%
  mutate(time_gain = ArrDelay - DepDelay, speed = Distance/AirTime *60)
calc
```

to summarise and sum an individual column

```{r}
View(summarise(flights,Delay =sum(DepDelay, na.rm = TRUE))
```