---
title: "Lecture 4"
author: "Joshua Matthew(SCV2021)"
date: "24/10/2019"
output: pdf_document
---
import the diamond data
```{r}
dmd<- read.csv("data/diamond.csv", header = T, dec = '.', sep = ',')
```
libraries that are needed
```{r}
library(tidyverse)
library(readr)
library(readxl)
library(ggplot2)
```
using ggplot2 on mbta data
```{r}
New_dta<-read_excel("data/mbta.xlsx", skip = 1, range =  cell_cols(2:60))
```
a peek into the diamond data
```{r}
glimpse(dmd)
```

plot carat and price; aes is the aestethics.

```{r}
ggplot(dmd, aes(x = carat, y = price))
```
adding the the geom functions (point,bar, boxplot,jitter)
```{r}
ggplot(dmd, aes(x = carat, y = price))+geom_point()
```
gathering the year into a column against passengers
```{r}
gathered_New_dta<- New_dta%>%gather(`2007-01`: `2011-10`, key = 'Year', value = 'Passengers')
View(gathered_New_dta)
```



```{r}
tidy_mbta<-gathered_New_dta%>%separate(Year, into = c("Year", "Month"))
View(tidy_mbta)
```


```{r}
filtered_New_dta<- filter(tidy_mbta ,mode %in% c("Commuter Rail", "Heavy Rail", "Light Rail", "Private Bus"))
View(filtered_New_dta)
```


convert data to factor(fct), numeric and any form
```{r}
glimpse(filtered_New_dta)
```
converting the reference to mode, year, month and passengers into factor in order to suit ggplot...factors are more repititive and numeric is discrete.

```{r}
filtered_New_dta$mode<- as.factor(filtered_New_dta$mode)
filtered_New_dta$Year<- as.factor(filtered_New_dta$Year)
filtered_New_dta$Passengers<- as.numeric(filtered_New_dta$Passengers)
filtered_New_dta$Month<- as.factor(filtered_New_dta$Month)
```



```{r}
glimpse(filtered_New_dta)
```

using ggplot to view the filtered_New_dta data of passengers against Year and colour it based on mode or month (factor /categorical type)

```{r}
filtered_plot<-ggplot(filtered_New_dta,aes(x = Year, y =Passengers)) + geom_point(aes(color =mode))
```


```{r}
p1<-filtered_plot+labs(title = "Passengers against Year", x= "Year dah dah dah", y = "Passengers")
```

```{r}
p1
```

```{r}
diamonds<-ggplot(dmd, aes(x = carat, y = price))
diamonds+ geom_point(aes(colour = color, shape= cut, size = 0.6))+labs(title = "Price against carat", x= "Carat", y= "Price")
```

```{r}
diamondsPlot<-diamonds+ geom_point(aes(colour = color, shape= cut, size = 0.6))+labs(title = "Price against carat", x= "Carat", y= "Price")+
  scale_x_continuous(limits = c(0,4), breaks = seq(0,4, by= 0.5))+
  scale_y_continuous(limits = c(0,1600), breaks = seq(0,1600, by = 2000))
suppressWarnings(print(diamondsPlot))

```

```{r}

```
```{r}

```
```{r}

```
```{r}

```

```{r}

```


