---
title: "Lecture2"
author: "Josh"
date: "10/10/2019"
output: pdf_document
---

#use choose.file() to load a folder and select a specific file  or use data/file_path to load the file directly into rmd
```{r}
who <- read.table("data/who_suicide_statistics.csv",header = T, sep = ",")
who2 <- read.table("data/who_suicide_statistics_semi.csv",header = T, sep = ";")


```

```{r}
autoData <- read.table("data/auto.txt",header = T, sep = ",")
```

```{r}
View(who)
View(who2)
```


```{r}
names(autoData) <- c("sym","norm","make","f_type","asp","num_doors","body_style","drive_wheel","engine_loc","wheel_base","length","width","height","curb_weight","engine_type","num_cyl","eng_size","fuel_type","bore","stroke","comp","horse","peak_rpm","city_mpg","highway","price")
#naming the columns (header)
# using $ for referencing
```

```{r}
View(autoData)
```

```{r}
length(autoData)
```

```{r}
potato1 <- read.table("data/potatoes.txt",header = T, sep = "\t")
```

```{r}
View(potato1)
```

#load the package  readr for csv by using read_extension
```{r}
library(readr)
```

```{r}
home<- read_csv(file.choose(), col_names = T)
```

```{r}
View(home)
```

#load the package to read excel files
```{r}
library(readxl)
```
to read an excel file in distinct sheets
```{r}

excel1<- read_excel("data/Employees.xlsx", sheet = 1)
```

```{r}
excel2<- read_excel("data/Employees.xlsx", sheet = 2)

```

```{r}
View(excel1)
```

```{r}
excel2
```

using row function 'head'(first 6) and 'tail'(last 6 rows)
also specifying distinct columns using vectors i.e c()
```{r}
head(autodata[,1:4])
```

```{r}
tail(autoData[, c(1,3,5)])
```

loading the DBI for database connection
```{r}
library(DBI)
library(RMySQL)
```

connecting the cloud host through R

```{r}
connect<- dbConnect(RMySQL::MySQL(), dbname = "tweater", host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", port = 3306, user = 'student', password = 'datacamp')
```


connecting MySQL local host to R
```{r}
cape_codd<- dbConnect(RMySQL::MySQL(), dbname = "cape_codd", host = "127.0.0.1", port = 3306, user = 'root', password = 'Adeniyi96@123')
```


store all tables in the connected database in variable 'tables'
```{r}
tables<- dbListTables(connect)
```
to view  a list of all tables in the database 'connect'

```{r}
tables
```
to select distinct table into a specified variable
```{r}
user_tables<- dbReadTable(connect,"users")

```

to view user_tables
```{r}
View(user_tables)
```
to extract 'comment' table from 'tables'
```{r}
user_comments<- dbReadTable(connect,"comments")
users<-dbReadTable(connect, "users")
```



to view users also user_comments
```{r}
users
View(users)
user_comments

```
to export or save a table in R as a csv file
```{r}
write.csv(user_comments, file = "userComments.csv")
```
to save as txt file or csv
```{r}
write.table(users, file = "users.txt")
write.csv(users, file = "users.csv")
```

```{r}
houses<-read.csv("https://factual.ng/training/house.csv", header = T)
```

```{r}
View(houses)
```
using package 'haven'
```{r}
library(haven)
```
saving a dta (stata file) 'alcohol' as 'd_ta'
```{r}
d_ta<- read_dta("data/alcohol.dta")
```
viewing the saved d_ta file
```{r}
View(d_ta)

```
using 'str' to view its structure
```{r}
str(d_ta)
```

to load csv file (house) as 'scat' for scattered plots; used for continuous variables not categorical varibles.
```{r}
scat<-read.csv("data/house.csv", header = T)
```


```{r}
View(scat)
```
to plot two variables, using the '$' and formatting the plot with 'main', 'ylab', 'xlab'
```{r}

BathroomPrice<-plot(scat$Bathrooms, scat$Price, main = "Relationship btw Bathroom and Price",xlab = "Bathroom",ylab = "Pricing")
write.csv(BathroomPrice, file = "bathroom")
```

```{r}
hist( log(scat$Price),xlab = "Price", main = "Histogram of Price")
```


```{r}
hist(scat$SQFT, xlab = "Location", main = "Histogram of Price")
```




Assignment:
Load the diamond data set into R. Plot the scatter plot of the variables "price" and "depth", with Price on the y-axis and depth on the x-axis.
Format your plot by putting title on the plot.
Label your X-axis and Y-axis as well.

carry out histogram of the "price" variable and the Title your plots accordingly.

hamid@factual.ng
```{r}
``